Museum specimens as baseline for assessing a potential global pollinator crisis

### 2018-02-05 ms.Rmd
Add more ideas and figures

Define trends: Decline, turnover (changes in evennes winners loosers), exotics... Review current methods to assess 1) space for time (recent reviews rae, ...), 2) monitoring schemes /citicen science (iNatralis) (UK) and 3) museums type data. Highlight its importance (cite Primak). Pros and cons with an enphasis on Museums.

##Current evidence:
- US: Bartomeus 2013 and Burkle 201?, ... NAS book Critizize Burkle dessign and compare to mine.
- EU: Beijsmier, Carvalho, Schepper, ... UK? Peter Keil flies? Moths? Buterflies!
- South Africa: Pauw, maybe with plants...
- Specific examples e.g. Bombus dalbhomi (Aizen), ... Thomson ecology letters
 Fig 1: a global map showing evidence (% decline in richness) and Data deficient areas would be great.

##Other uses of museums in pollinators. Identifying mechanisms of decline
- Oliveira plos one 
- Osorio Ecol Lettres
- Phenoloy (Bartomeus)

>>>>

Define trends: Decline, turnover (changes in evennes winners loosers), exotics... Review current methods to assess 1) space for time (recent reviews rae, ...), 2) monitoring schemes /citicen science (iNaturalis) (UK) and 3) museums type data. Highlight its importance (cite Primak). Pros and cons with an enphasis on Museums.

##Current evidence:
- US: Bartomeus 2013 and Burkle 201?, ... NAS book Critizize Burkle dessign and compare to mine.
- EU: Beijsmier, Carvalho, Schepper, ... UK? Peter Keil flies? Moths? Buterflies!
- South Africa: Pauw, maybe with plants...
- Specific examples e.g. Bombus dalbhomi (Aizen), ... Thomson ecology letters

We compiled current evidence worldwide and show that ... (short methods included in text)

```{r map, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(rworldmap)
map <- data.frame(iso2 = c("ES", "NZ"), richness = c(88, 75))
#FAKE data, we need to compile a bbdd with the data at hand.
map2 <- joinCountryData2Map(map, joinCode="ISO2", nameJoinColumn = "iso2", nameCountryColumn = "distribution")
mapCountryData(mapToPlot = map2, numCats = 2, nameColumnToPlot = "richness"
               , mapTitle = "Bee richness decline", addLegend = FALSE)

```

Fig 1: a global map showing evidence (% decline in richness) and Data deficient areas would be great.

## Beyond species occurrences
Other uses of museums in pollinators. Identifying mechanisms of decline
- Oliveira plos one 
- Osorio Ecol Lettres
- Phenoloy (Bartomeus)

## Prioritizing the low hanging fruit.

Analyze trade off between number of records in Gbif for > 1980 and pre <1980. Each quadrant needs an analysis. 
+ we provide code to run it, and maybe add it to the repo per country? distributed database style?
We do it for two regions:

```{r low hanging fruits, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(rgbif)
#spain_code <- isocodes[grep("Spain", isocodes$name), "code"]
#portugal_code <- isocodes[grep("Portugal", isocodes$name), "code"]
apidae_key <- name_backbone(name="Apidae", rank = "family")$usageKey
andrenidae_key <- name_backbone(name="Andrenidae", rank = "family")$usageKey
halictidae_key <- name_backbone(name="Halictidae", rank = "family")$usageKey
colletidae_key <- name_backbone(name="Colletidae", rank = "family")$usageKey
megachilidae_key <- name_backbone(name="Megachilidae", rank = "family")$usageKey
stenotritidae_key <- name_backbone(name="Stenotritidae", rank = "family")$usageKey
melittidae_key <- name_backbone(name="Melittidae", rank = "family")$usageKey

keys <- c(apidae_key, andrenidae_key, halictidae_key, halictidae_key, colletidae_key,
          megachilidae_key, stenotritidae_key, melittidae_key)
#loop through all families, through all countries and store in datarame
country_codes <- c("ES", "NZ", "PT", "GE") #need to fetch them all EASY using rgbif
lhf <- matrix(ncol = 1+(7*2), nrow = length(country_codes))
lhf <- as.data.frame(lhf)
colnames(lhf) <- c("country", "apidaepre1980", "andrenidaepre1980", "halictidaepre1980", "colletidaepre1980", 
                    "megachilidaepre1980", "stenotritidaepre1980", "melittidaepre1980",
                   "apidaepost1980", "andrenidaepost1980", "halictidaepost1980", "colletidaepost1980", 
                    "megachilidaepost1980", "stenotritidaepost1980", "melittidaepost1980")
for(i in 1:length(country_codes)){
    for(j in 1:length(keys)){
        lhf[i,1] <- country_codes[i]
        lhf[i,1+j] <- as.numeric(occ_count(taxonKey= keys[j],
                            from = 1980, to = 2018,
                            georeferenced=TRUE,
                            country = country_codes[i]))
        lhf[i,1+j+7] <- as.numeric(occ_count(taxonKey= keys[j],
                            from = 1850, to = 1980,
                            georeferenced=TRUE,
                            country = country_codes[i]))
    }
}
#lhf
plot(rowSums(lhf[,2:8]), rowSums(lhf[,9:15]), xlab = "post1980", ylab = "pre1980", col = as.factor(lhf$country))
abline(h = 3000, v = 5000)
#plot needs legend by country, and to be ggplotted

```

Fig 2: number of bee occurrences pre 1980 and post 1980 in gbif. Upper rigth cuadrat are well covered countries, upper left cuadrat are countries that need resurvey of old sites. Lower right are sites which needs to digg museum records and lower left are countries hard to get data.

We ilustrate how this can be done using two countries and providing code to reproduce our analysis, so that more countries can contribute.

- Describe data and analysis.
- Present main results

- Describe data and analysis.
- Present main results
